---
title: "Example data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Included example data sets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<style>
body {
text-align: justify}
</style>
```{r, echo = FALSE, eval=TRUE, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(knitr.table.format = "html", rmarkdown.html_vignette.check_title = FALSE)

library(eRTG3D)
library(raster)
library(sp)
set.seed(123)
```

## Bird trajectory
An example GPS track of a stork `niclas` and an example Digital Elevation Model `dem` are included in the package. The DEM is a snippet from the *SRTM 90m Digital Elevation Database v4.1* from [CGIAR-CSI](http://srtm.csi.cgiar.org/srtmdata/) and the stork track is a subset of the data in the study *LifeTrack White Stork SWGermany 2014-2017* (Movebank Study ID 21231406) from the [Movebank Data Repository](https://www.datarepository.movebank.org). 

The area covered by the test data is the Swiss midland. The example track is a data.frame with coordinates in x, y and z. The x and y coordinates are in the CH1903+ / LV95 (EPSG:2056) reference system.
```{r }
head(niclas)
```

## Digital elevation model
The DEM is a rasterLayer of the raster package, it's Coordinate Reference System (CRS) is also CH1903+ / LV95. The stork flies from southwest to northeast, GPS fixes are collected every 15 minutes.

```{r }
dem
```
